---
output: github_document
---

<!-- README.md is generated from dev/README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
# set options
Sys.setenv(OPENSSL_CONF = "/dev/null")
options(talib.chart.modebar = FALSE)

knitr::opts_knit$set(
	base.dir = normalizePath("..")
)

knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>",
	fig.path = "man/figures/README-",
	fig.dpi = 120,
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)
```

# {talib}: Fast TA-Lib indicators and candlestick patterns for R <a href="https://serkor1.github.io/ta-lib-R/"><img src="man/figures/logo.svg" align="right" height="170" alt="talib website" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/serkor1/ta-lib-R/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/serkor1/ta-lib-R/actions/workflows/R-CMD-check.yaml)
[![Remote Install](https://github.com/serkor1/ta-lib-R/actions/workflows/remote-install.yaml/badge.svg)](https://github.com/serkor1/ta-lib-R/actions/workflows/remote-install.yaml)
[![Codecov test coverage](https://codecov.io/gh/serkor1/ta-lib-R/graph/badge.svg)](https://app.codecov.io/gh/serkor1/ta-lib-R)
[![CRAN status](https://www.r-pkg.org/badges/version/talib)](https://CRAN.R-project.org/package=talib)
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/last-month/talib?color=blue)](https://r-pkg.org/pkg/talib)
<!-- badges: end -->

[{talib}](https://serkor1.github.io/ta-lib-R/) provides fast R bindings
to the [TA-Lib](https://github.com/TA-Lib/ta-lib) C library for OHLCV data: technical indicators, candlestick pattern recognition, rolling-window utilities, and composable financial charts. 
It is designed for researchers, analysts, and quant developers who need technical-analysis features in R without building a heavy dependency stack. Core computations are executed in C through `.Call()`, while charting support is available through optional [{plotly}](https://github.com/plotly/plotly.R) and [{ggplot2}](https://ggplot2.tidyverse.org/) integrations.[^1]

The API covers 150+ [TA-Lib](https://github.com/TA-Lib/ta-lib)-backed functions across momentum, overlap, volatility, volume, cycle, price-transform, rolling-statistics, and candlestick-pattern families, including 61 candlestick pattern detectors.

## Why {talib}?

<div align="center">

| Need | {talib}|
|:----------------------|:-----------------------------------------------------------------------------------------|
| Technical indicators | TA-Lib-backed moving averages, momentum, volatility, volume, cycle, and overlap studies   |
| Candlestick patterns | Built-in Japanese candlestick pattern recognition                                         |
| OHLCV workflows      | Works directly with open, high, low, close, and volume columns                            |
| Performance          | Computation delegated to C routines through `.Call()`                                     |
| Charts               | Composable financial charts with optional `{plotly}` and `{ggplot2}` support              |

</div>

## Installation[^2]

Install the release version from CRAN:

```{r install-cran, echo = TRUE, eval = FALSE}
install.packages("talib")
```

Install the development version from GitHub:

```{r install-dev, echo = TRUE, eval = FALSE}
pak::pak("serkor1/ta-lib-R")
```

## Quick start

All functions provide S3 methods for `<xts>`, `<data.frame>`, `<matrix>`, and---where applicable---`<vector>` inputs. The general convention is simple: the output uses the same container type as the input.

```{r quick-start, echo=TRUE}
## calculate the
## relative strength index
relative_strength_index <- talib::RSI(
	talib::GOOGL
)

## check class equivalence
inherits(
	relative_strength_index, 
	class(talib::GOOGL)
)

## display results
tail(
	relative_strength_index
)
```

Indicator outputs preserve input length, which keeps results aligned with the
original OHLCV rows.

```{r multiple-indicators, echo = TRUE}
## combine multiple
## indicators
features <- cbind(
    talib::relative_strength_index(talib::GOOGL),
    talib::bollinger_bands(talib::GOOGL),
    talib::engulfing(talib::GOOGL)
)

tail(features)
```

## Charting 

[{talib}](https://serkor1.github.io/ta-lib-R/) comes with a composable charting API built on two core functions: `indicator()` and `chart()`---both functions are built on `model.frame` for maximum flexibility:

```{r charting-example, echo=TRUE}
## subset data and
## store as 'GOOGL'
GOOGL <- talib::GOOGL[1:75, ]

## construct chart in a brace block
## alternatively use `|>`
{
	## initialize main chart
    talib::chart(
		x     = GOOGL,
		title = "Alphabet Inc."
	)

	## add Bollinger Bands to
	## the existing chart
    talib::indicator(
		talib::BBANDS
	)

	## add Simple Moving Averages (SMA)
	## to the chart in a loop
	for (timePeriod in seq(5, 15, by = 3)) {
		talib::indicator(
			talib::SMA,
			timePeriod = timePeriod
		)
	}

	## similar subchart indicators
	## like the Relative Strength Index
	## can be grouped to avoid repeated
	## subpanels
    talib::indicator(
        talib::RSI(timePeriod = 10),
        talib::RSI(timePeriod = 14),
        talib::RSI(timePeriod = 21)
    )

	## identify Doji patterns
	## and add them to the chart
	talib::indicator(
		talib::doji
	)
}
```

## Implementation: {talib} vs upstream (TA-Lib Core)

Functions use descriptive snake_case names; each is aliased to its [TA-Lib](https://github.com/TA-Lib/ta-lib) shorthand for compatibility with the broader ecosystem, and to a camelCase name for consistency across R's finance ecosystem:

<div align="center">

| Category              | TA-Lib (C)           | {talib}                     | {talib} alias       | {talib} camelCase alias     |
|:----------------------|:---------------------|:----------------------------|:--------------------|:----------------------------|
| Overlap Studies       | `TA_BBANDS()`        | `bollinger_bands()`         | `BBANDS()`          | `bollingerBands()`          |
| Momentum Indicators   | `TA_CCI()`           | `commodity_channel_index()` | `CCI()`             | `commodityChannelIndex()`   |
| Volume Indicators     | `TA_OBV()`           | `on_balance_volume()`       | `OBV()`             | `onBalanceVolume()`         |
| Volatility Indicators | `TA_ATR()`           | `average_true_range()`      | `ATR()`             | `averageTrueRange()`        |
| Price Transform       | `TA_AVGPRICE()`      | `average_price()`           | `AVGPRICE()`        | `averagePrice()`            |
| Cycle Indicators      | `TA_HT_SINE()`       | `sine_wave()`               | `HT_SINE()`         | `sineWave()`                |
| Pattern Recognition   | `TA_CDLHANGINGMAN()` | `hanging_man()`             | `CDLHANGINGMAN()`   | `hangingMan()`              |

</div>

### Interface: R vs Python

The main difference between the R and Python interfaces is how OHLCV series are passed into each indicator function. 
Below is an example of identifying `Doji` patterns in R and Python.

In Python, each series is passed independently:

```python
import numpy as np
import talib

o = np.array([1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1], dtype=float)
h = np.array([2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2], dtype=float)
l = np.array([1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1], dtype=float)
c = np.array([2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 1], dtype=float)

print(
	talib.CDLDOJI(o, h, l, c)
)
```

In R the series are passed as a tabular container:

```{r r-vs-python-block, echo = TRUE, eval = FALSE}
ohlc <- data.frame(
	open  = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
	high  = c(2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2),
	low   = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
	close = c(2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 1) 
)

talib::CDLDOJI(
	ohlc
)
```

All default series arguments are handled internally, and the `R` interface is therefore higher-level: users pass one OHLC container rather than manually splitting the series.

## Contributing and cloning

Contributions are welcome. For non-trivial changes, please open an issue first to discuss the proposed design, API impact, and testing approach.

This repository vendors [TA-Lib](https://github.com/TA-Lib/ta-lib) as a Git submodule. Clone the repository with submodules enabled:

```sh
git clone --recurse-submodules https://github.com/serkor1/ta-lib-R.git
cd ta-lib-R
```

If you already cloned the repository without submodules, initialize them with:

```sh
git submodule update --init --recursive
```

Most indicator wrappers, helper functions, documentation fragments, and unit tests are generated from the scripts in `codegen/`. The charting interface is maintained separately.

Common development tasks are exposed through Make targets:

```sh
make help
```

See CONTRIBUTING.md for the full development workflow.

## Code of Conduct

Please note that [{talib}](https://serkor1.github.io/ta-lib-R/) is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

[^1]: See `benchmark/` for detailed benchmarks against [{TTR}]() and general performance across multiple indicators.
[^2]: `{talib}` is a compiled package. CRAN binaries are available for standard platforms when provided by CRAN. Source installation requires a working compiler toolchain and [CMake](https://cmake.org/).