---
title: "{talib}: R bindings to TA-Lib"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{{talib}: R bindings to TA-Lib}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
## set chart options
options(talib.chart.dark = FALSE)

knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>",
	out.width = "100%",
	out.height = "680",
	fig.align = "center"
)
```

[{talib}](https://github.com/serkor1/ta-lib-R/) is an R package for Technical Analysis (TA) and interactive financial charts. The package is a wrapper of [TA-Lib](https://github.com/TA-Lib/ta-lib) and supports over 200 indicators, including candlestick patterns.

## Basic usage

In this section I will show how to calculate and chart different indicators using Bitcoin (BTC). BTC is a built-in `data.frame` using the [{cryptoQuotes}](https://github.com/serkor1/cryptoQuotes)-package and has the following form:

```{r}
str(
	talib::BTC
)
```

The focus of this section will be Bollinger Bands and Harami patterns. All remaining indicators and candlestick patterns behaves in the same manner.

> **NOTE:** All functions expects that a permutation of either `open`, `high`, `low`, `close` and/or `volume` exists (case sensitive). But this depends on the specific functions; for indicators that only uses `high` and `low`, the remaining columns needs not to be present in the data. Column names must follow OHLC-V conventions unless remapped via the `cols`-argument.

### Indicators

All indicator functions in [{talib}](https://github.com/serkor1/ta-lib-R/) assumes that the input argument `x` is an OHLC-V object coercible to `data.frame`; the `BTC` can passed directly into the function:  

```{r Bollinger Bands (default)}
{
	cat("Bollinger Bands")
	tail(
		talib::bollinger_bands(
			x = talib::BTC
		)
	)
}
```

Internally, `bollinger_bands()` calls `model.frame` with a prespecified `formula` corresponding to the default calculation method of the indicator. You can modify this by passing a different variable through the `cols`-argument as follows:

```{r Bollinger Bands (custom)}
{
	cat("Bollinger Bands")
	tail(
		talib::bollinger_bands(
			x = talib::BTC,
			cols = ~high
		)
	)
}
```

The Harami pattern, and all candlestick patterns in general, returns a `matrix` of \<integers\>: -1 for bearish patterns, 1 for bullish pattern and 0 for no pattern. See below:

```{r Harami}
{
	cat("Harami Patterns")
	tail(
		talib::harami(
			talib::BTC
		)
	)
}
```

The pattern is rare, so let us find occurrences instead:

```{r Harami(assigned)}
## assign the Harami
## pattern
x <- talib::harami(
	talib::BTC
)
```

```{r Harami(Bullish)}
## locate bullish Harami
## patterns
{
	cat("Bullish Harami")
	talib::BTC[
		which(x == 1),
	]
}
```

```{r Harami(Bearish)}
## locate bearish Harami
## patterns
{
	cat("Bearish Harami")
	talib::BTC[
		which(x == -1),
	]
}
```

### Charting indicators

In this section I will focus on the basics of charting. The charting uses [{plotly}](https://github.com/plotly/plotly.R) internally which also enables manual drawing of support and resistance lines.

> **Information:** Please refer to `vignette(topic = "charting", package = "talib")`
> for a more detailed description of the charting.

The charting functionality mimicks the behaviour of `plot()` and `lines()` where each is called separately. There are two main charting functions in [{talib}](https://github.com/serkor1/ta-lib-R/): `chart()` and `indicator()`. Below I will expand the analysis by adding Moving Average Convergence/Divergence indicators to the chart:

```{r chart(basic)}
{
	## main candlestick
	## chart
	talib::chart(
		talib::BTC
	)

	## add bollinger bands
	## to the chart
	talib::indicator(
		talib::bollinger_bands
	)

	## add identified harami
	## patterns
	talib::indicator(
		talib::harami
	)

	## add moving average
	## convergence/divergence
	## to the chart
	talib::indicator(
		talib::moving_average_convergence_divergence
	)
}
```

One thing to note here is that `indicator()` internally calls the [{plotly}](https://github.com/plotly/plotly.R)-methods for each indicator which determines whether the indicator is main-chart indicator or sub-chart indicator.

## Contributing and bug reports

The underlying library, [TA-Lib](https://github.com/TA-Lib/ta-lib), is a long-standing and well-known library but this R wrapper is still in its early stage. All contributions, suggestions, and bug reports are welcome.