---
title: "{talib}: R bindings to TA-Lib"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{talib}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
## set chart options
options(talib.chart.dark = FALSE)

knitr::opts_chunk$set(
  collapse      = TRUE,
  comment       = "#>",
  out.width     = "100%",
  out.height    = "680",
  fig.align     = "center"
)
```

```{r setup}
library(talib)
```

[{talib}](https://github.com/serkor1/curly-giggle) is wrapper of the 'C' library [TA-Lib](https://github.com/TA-Lib/ta-lib). 
The library supports indicators for Technical Analysis, including algorithmic candlestick pattern recognition.

The implementation has been extended to include interactive financial charts with a [{plotly}](https://github.com/plotly)-backend. All indicators in [{talib}](https://github.com/serkor1/curly-giggle)
are soft-wrappers of `model.frame` for added flexibility.

## Basic usage

### Simple Moving Averages

Below is an example on calculating the simple moving average of the Bitcoin (BTC):

```{r SMA(simple)}
tail(
  talib::SMA(
    x = talib::BTC
  )
)
```

The simple moving average can be calculated over a specified range via `...` which gets forwarded to `model.frame`, and calculated for multiple prices, as follows:

```{r SMA(flexible)}
tail(
  talib::SMA(
    x = talib::BTC,
    cols = ~low + high,
    ## calculate simple moving average
    ## up and until the 100th indices
    subset = 1:nrow(talib::BTC) <= 100
  )
)
```

### Charting indicators

> **Information:** Please refer to the `charting` vignette
> for a more detailed description of the charting.

The charting functionality mimicks the behaviour of `plot()` and `lines()` where each is called separately. 

```{r chart(basic)}
{
  ## candlestick chart
  talib::chart(
    talib::BTC
  )

  ## add SMA
  talib::indicator(
  talib::SMA(),
  cols = ~high + low
  )

  ## add acceleration bands
  talib::indicator(
    talib::ACCBANDS()
  )
}

```
